---
title: An R Markdown document converted from "notebook/ms2deepscore/tutorial.ipynb"
output: html_document
---

# MS2DeepScore API – Quickstart

*In order to use omigami, you need an access token provided by DataRevenue.*
## Introduction

This is a short guide on how you can predict molecular structural similarities (Tanimoto scores) from pairs of mass spectrometry spectra in local mgf files - using Omigami's MS2DeepScore API.

The API uses a MS2DeepScore model which was trained on > 100,000 MS/MS spectra from GNPS. It prepares the two spectra and using a Siamese neural network predicts returns the molecular structural similarity (Tanimoto score).

This notebook shows you how to:
1. Specify any MGF file with two spectra.
2. Make a prediction through the MS2DeepScore API.

# 1. Load a MS/MS dataset (MGF format)

We'll use the MS/MS dataset in the MGF format in this directory. You can also select any other from the [GNPS spectral library](https://gnps-external.ucsd.edu/gnpslibrary), or - of course - use your own.

*Note that MS2DeepScore accepts two spectra at a time. Please prepare a MGF file with only two spectra.*

```{r}
# Load your own MS/MS dataset
path_to_mgf <- 'GNPS-2-SPECTRA.mgf'
```

# 2. Predict molecular structural similarity with MS2DeepScore

`Romigami` is a R wrapper which instantiates a virtual environment on which to install and call the Python `Omigami` package

`MS2DeepScore` is a python wrapper which :
- Builds a json payload from the MGF file
- Calls the MS2DeepScores API
- Formats the prediction results into readable format

```{r}
# install devtools and romigami if not already installed
if("devtools" %in% installed.packages() == FALSE) {install.packages("devtools")}
devtools::install_github("omigami/romigami")

# import the romigami package
library("romigami")
```

```{r}
# Initialize omigami environment
omigami <- omigami_init()
```

```{r}
# Run Spec2Vec library search with your user token
client <- omigami$MS2DeepScore(token="my_token")

n_best_matches <- 10
include_metadata <- list("Smiles", "Compound_name")
ion_mode <- "positive"  # either positive or negative
results <- client$match_spectra_from_path(mgf_path = path_to_mgf,
                                          n_best = n_best_matches,
                                          include_metadata = include_metadata,
                                          ion_mode = ion_mode
            )
```

## 2.1 View prediction
A dictionary is returned.

```{r}
results
```

____

